Analysis of Open Data and Computational Reproducibility in Registered Reports in Psychology

### 2019-05-23 reproducing_registered_reports.Rmd
Update reproducing_registered_reports.Rmd
* Comma in abstract for clarity
* MATLAB is a capitalized trademark

abstract: |
  Ongoing technological developments have made it easier than ever before for scientists to share their data, materials, and analysis code. Sharing data and analysis code makes it easier for other researchers to re-use or check published research. These benefits will only emerge if researchers can reproduce the analysis reported in published articles, and if data is annotated well enough so that it is clear what all variables mean. Because most researchers have not been trained in computational reproducibility, it is important to evaluate current practices to identify practices that can be improved. We examined data and code sharing, as well as computational reproducibility of the main results without contacting the original authors, for Registered Reports published in the in psychological literature between 2014 and 2018. Of the 62 articles that met our inclusion criteria data was available for 40 articles, and analysis scripts for 43 articles. For the 35 articles that shared both data and code and performed analyses in SPSS, R, or JASP, we could run the scripts for 30 articles, and reproduce the main results for 19 articles. Although the percentage of articles that shared both data and code (61%) and articles that could be computationally reproduced (54%) was relatively high compared to other studies, there is clear room for improvement. We provide practical recommendations based on our observations, and link to examples of good research practices in the papers we reproduced.

@@

To find Registered Reports published in psychology we drew from a database of registered reports maintained by the Center for Open Science.^[https://www.zotero.org/groups/479248/osf/items/collectionKey/KEJP68G9] At the start of this project (06-05-2018) this database consisted of 118 published Registered Reports. Eighty-one articles in this database were published in the psychological literature. We limited our analysis to studies performed by single groups, since such articles are most representative of the researchers currently work, and excluded 7 large scale collaborations where dedicated team members were responsible for making the analysis reproducible, such as Registered Replication Reports [@hagger_multilab_2016] and Many Labs projects [@klein_investigating_2014; @klein_many_2018]. Upon further inspection of the 72 papers left in the dataset, it turned out that 10 were not formally Registered Reports. This left `r length(all_data$doi)` studies in our sample. When evaluating whether we could reproduce the original results, we limited ourselves to statistical software packages that we had experience with (R, SPSS, and JASP) and excluded studies that required expertise in software packages we are not trained in (e.g., dedicated EEG software, Python, Matlab), which led to the exclusion of `r sum(all_data$included_language == FALSE, na.rm = TRUE)` additional studies.

@@

We also coded the extent to which data and code was specific to software that was not freely available. When open source software is used, the analyses can be reproduced by anyone with a computer and internet, regardless of whether they have bought the same software package. When proprietary software is used, results might still be reproducible by anyone, but it could require more effort to do so. For example, SPSS produces proprietary .sav and .sps files. However, .sav files can be opened in R, and .sps files can be opened by a text editor and the code can be rewritten, as long the code is annotated well enough to be recoded into R. Note that we had access to SPSS and therefore reproduced the analyses using the original .sps scripts where available. The data files of one article where an EEG study was reported consisted of .eeg, .vhdr, and vmrk files that were analyzed in matlab, which we could not reproduce using open source software. This left `r sum(all_data$not_software_specific, na.rm = TRUE)` papers where data was available in a non-software-specific format.

@@

For each article two coders examined whether the code could be run on the data. Of the original `r length(all_data$doi)` articles, `r sum(all_data$included_language == FALSE, na.rm = TRUE, na.rm = TRUE)` used Matlab and Python and were excluded from the analysis due to a lack of expertise in our team with these languages. In total code and data were available for `r sum(all_data$analysis_script_included ==TRUE & all_data$availability == TRUE, na.rm = TRUE)` articles, of which `r sum(all_data$analysis_script_included ==TRUE & all_data$availability == TRUE & all_data$included_language ==TRUE, na.rm = TRUE)` used R, SPSS, or JASP. We were able to run the code for `r sum(all_data$run_script_final == TRUE, na.rm = TRUE)` out of the `r sum(all_data$analysis_script_included ==TRUE & all_data$availability == TRUE & all_data$included_language ==TRUE, na.rm = TRUE)` articles that used SPSS, R, and JASP. R was used as a coding language in `r length(run_script_r)`, of which `r sum(run_script_r)` scripts could be run on the data, `r length(run_script_spss)` papers used SPSS, for `r sum(run_script_spss)` of which the script could be run, and `r length(run_script_r_spss)` papers used both SPSS and R, for which the scripts could be run for all `r sum(run_script_r_spss)` papers. JASP does not seperate the data and code, but instead stores both in a single JASP file. This means that the analyses are always directly linked to the code for every output. It is always clear which settings were used to generate results. Because of this useful feature, the `r sum(run_script_jasp)` articles that relied on JASP for the analyses always allowed us to reproduce the analyses.

>>>>

abstract: |
  Ongoing technological developments have made it easier than ever before for scientists to share their data, materials, and analysis code. Sharing data and analysis code makes it easier for other researchers to re-use or check published research. These benefits will only emerge if researchers can reproduce the analysis reported in published articles, and if data is annotated well enough so that it is clear what all variables mean. Because most researchers have not been trained in computational reproducibility, it is important to evaluate current practices to identify practices that can be improved. We examined data and code sharing, as well as computational reproducibility of the main results without contacting the original authors, for Registered Reports published in the in psychological literature between 2014 and 2018. Of the 62 articles that met our inclusion criteria, data was available for 40 articles, and analysis scripts for 43 articles. For the 35 articles that shared both data and code and performed analyses in SPSS, R, or JASP, we could run the scripts for 30 articles, and reproduce the main results for 19 articles. Although the percentage of articles that shared both data and code (61%) and articles that could be computationally reproduced (54%) was relatively high compared to other studies, there is clear room for improvement. We provide practical recommendations based on our observations, and link to examples of good research practices in the papers we reproduced.

@@

To find Registered Reports published in psychology we drew from a database of registered reports maintained by the Center for Open Science.^[https://www.zotero.org/groups/479248/osf/items/collectionKey/KEJP68G9] At the start of this project (06-05-2018) this database consisted of 118 published Registered Reports. Eighty-one articles in this database were published in the psychological literature. We limited our analysis to studies performed by single groups, since such articles are most representative of the researchers currently work, and excluded 7 large scale collaborations where dedicated team members were responsible for making the analysis reproducible, such as Registered Replication Reports [@hagger_multilab_2016] and Many Labs projects [@klein_investigating_2014; @klein_many_2018]. Upon further inspection of the 72 papers left in the dataset, it turned out that 10 were not formally Registered Reports. This left `r length(all_data$doi)` studies in our sample. When evaluating whether we could reproduce the original results, we limited ourselves to statistical software packages that we had experience with (R, SPSS, and JASP) and excluded studies that required expertise in software packages we are not trained in (e.g., dedicated EEG software, Python, MATLAB), which led to the exclusion of `r sum(all_data$included_language == FALSE, na.rm = TRUE)` additional studies.

@@

We also coded the extent to which data and code was specific to software that was not freely available. When open source software is used, the analyses can be reproduced by anyone with a computer and internet, regardless of whether they have bought the same software package. When proprietary software is used, results might still be reproducible by anyone, but it could require more effort to do so. For example, SPSS produces proprietary .sav and .sps files. However, .sav files can be opened in R, and .sps files can be opened by a text editor and the code can be rewritten, as long the code is annotated well enough to be recoded into R. Note that we had access to SPSS and therefore reproduced the analyses using the original .sps scripts where available. The data files of one article where an EEG study was reported consisted of .eeg, .vhdr, and vmrk files that were analyzed in 
, which we could not reproduce using open source software. This left `r sum(all_data$not_software_specific, na.rm = TRUE)` papers where data was available in a non-software-specific format.

@@

For each article two coders examined whether the code could be run on the data. Of the original `r length(all_data$doi)` articles, `r sum(all_data$included_language == FALSE, na.rm = TRUE, na.rm = TRUE)` used MATLAB and Python and were excluded from the analysis due to a lack of expertise in our team with these languages. In total code and data were available for `r sum(all_data$analysis_script_included ==TRUE & all_data$availability == TRUE, na.rm = TRUE)` articles, of which `r sum(all_data$analysis_script_included ==TRUE & all_data$availability == TRUE & all_data$included_language ==TRUE, na.rm = TRUE)` used R, SPSS, or JASP. We were able to run the code for `r sum(all_data$run_script_final == TRUE, na.rm = TRUE)` out of the `r sum(all_data$analysis_script_included ==TRUE & all_data$availability == TRUE & all_data$included_language ==TRUE, na.rm = TRUE)` articles that used SPSS, R, and JASP. R was used as a coding language in `r length(run_script_r)`, of which `r sum(run_script_r)` scripts could be run on the data, `r length(run_script_spss)` papers used SPSS, for `r sum(run_script_spss)` of which the script could be run, and `r length(run_script_r_spss)` papers used both SPSS and R, for which the scripts could be run for all `r sum(run_script_r_spss)` papers. JASP does not seperate the data and code, but instead stores both in a single JASP file. This means that the analyses are always directly linked to the code for every output. It is always clear which settings were used to generate results. Because of this useful feature, the `r sum(run_script_jasp)` articles that relied on JASP for the analyses always allowed us to reproduce the analyses.
